---
<<<<<<< HEAD
title: "Regular Expressions (Regexps)"
author: "Anna Deniz / Ángela Duarte"
output: 
  html_document:
    toc: TRUE
    df_print: paged
    number_sections: FALSE
    highlight: tango
    theme: lumen
    toc_depth: 3
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(knitr)
library(tidyverse)
library(stringr)
```

# Introduction

Regular Expressions, also known as Regexps, are conventions used to describe and find patterns in strings. They are useful not only in R but in other languages as well. 

“Strictly speaking, they are not so much a tool as they are a convention on how to query strings across a wide range of functions.” [Munzert et al., (Munzert et al., Automated Data Collection with R] (https://bit.ly/3BdbkhS). 

They may look *very* complicated if you are seeing them for the first time, but it is actually not that hard once you understand how they work. In this tutorial, we are going to show you how to apply them using `Stringr`, a package from the Tidyverse used for string manipulation. 

<<<<<<< HEAD
# Contents:
* What are Regexps good for?
* Key expressions
* How can you use Regexps to solve problems - Combinations
* Try it yourself - Exercises
* Where should you go to learn more
* Sources

---

# What are Regexps good for

Regular Expressions are very useful for retrieving information from a messy data set of strings. It is an alternative when manipulating the string to find the information you want is too much work. All you have to do is come up with a pattern for what you want to pull out and write it using Regexps’ syntax. 

Using Regular Expressions with `Stringr` is very convenient, as many of the functions in this package use 'pattern' as an input. 

Here is an example of how Regexps is used to get the right information from a messy list of phone numbers:

```{r}

phone_numbers <- c("(+49) 176 2692-5578", 
                   "My number is 176 2678 6789", 
                   "I don't have a phone", 
                   "176 4567-8923")

pattern <- ".*(\\d{3}).*(\\d{4}).*(\\d{4})"

str_match(phone_numbers, pattern)

```
Now, let's see how you can develop your own Regexps for your own messy data set!

---
# Key expressions

To demonstrate the key expressions, let's use this messy list of shopping items:

```{r}

shopping_list <- c("Apples - 5 green", "15 Oranges", " green grapes - 8 package", "2 packages of red grapes", "")

```

We will also use the function ‘str_match’ from the package Stringr, which receives as input the string and a pattern. In our case, the pattern arguments will be the Regular Expression.

MATCH CHARACTERS

1. The simplest Regular Expression you can write is the piece of string you want to match inside a quotation mark:

```{r}

pattern <- "grapes"

str_match(shopping_list, pattern)

```
2. \d matches the first digit in the string from 0 to 9. One thing to notice is that you have to add an extra \ before the expression for it to work, in order to "escape" the first \, that is a special character:
 
```{r}

pattern <- "\\d"

str_match(shopping_list, pattern)

```
3. \w matches the first alphanumeric element in the string (a-z, A-Z, 0–9). Here, you also have to use an extra \ to escape it:
 
```{r}

pattern <- "\\w"

str_match(shopping_list, pattern)

```
4. \s matches the first Whitespace (space, tab, newline) in the string. Again, don't forget to escape it:
 
```{r}

pattern <- "\\s"

str_match(shopping_list, pattern)

```
5. . matches any character (Wild card), apart from a void string:
 
```{r}

pattern <- "."

str_match(shopping_list, pattern)

```
QUANTIFIERS

6. + matches 1 or more of the matching patterns in the string that were specified before:
 
```{r}

pattern <- ".+"

str_match(shopping_list, pattern)

```
7. * matches 0 or more of the matching patterns in the string that were specified before, specially useful to find empty strings:
 
```{r}

pattern <- ".*"

str_match(shopping_list, pattern)

```
8. {n} will match exactly n of the matching patterns in the string that were specified before:

```{r}

pattern <- ".{3}"

str_match(shopping_list, pattern)

```
ALTERNATES

9. [ ] is used to match one of the characters inside of it:
 
```{r}

pattern <- "[lo5]"

str_match(shopping_list, pattern)

```
10. | is used to match one pattern or the other:
 
```{r}

pattern <- "Apple|grape"

str_match(shopping_list, pattern)

```
ANCHORS

11. ^ matches the beginning of a string:

```{r}

pattern <- "^."

str_match(shopping_list, pattern)

```
12. $ matches the end of a String:

```{r}

pattern <- ".$"

str_match(shopping_list, pattern)

```
GROUPS

13. ( ) is used to group the patterns inside of it:
 
```{r}

pattern <- "(grapes)"

str_match(shopping_list, pattern)

```
Now, let's practice more complicated, applied stuff!

---


# How can you use RegExp to solve problems? Using Combinations

Consider the following list of school objects:

```{r}
school_things=c("notebooks 450", "sheets 1000", "pencils 457", "colors 157", "nothing")
```

We will practice 5 combinations of the "Key Expressions" used above to do some interesting text manipulation. Let's start!

1. What if we wanted to get all elements from a certain kind in the `school_things` vector?

```{r}
str_match(school_things, "\\d+")
```
This `"\\d+"` means:"give me one or more digits from the vector `school_things`.

Similarly, we can use `"\\w+"` to get all the alpha numeric characters and `"\\s+"` to get all whitespaces. How would you translate them to plain English? 

```{r}
str_match(school_things, "\\w+") 
str_match(school_things, "\\s+") 
```


2. Separating numbers from the words. Hint: Use the expressions practiced in 1), `()` to group the strings and `\\W*` to remove all white spaces.  

```{r}
new_pattern="(\\w*)\\W*(\\d*)"
str_match(school_things, new_pattern)
```
Notice that the result is a matrix in which the columns are the vector itself, its alphanumeric values and its digits. 

3. How can we find the expressions that start with "not" on the vector? 
Notice that we will use `[[:alpha:]]` to get only the numerical part of it.

Let's learn the power of pattern matching!
```{r}

str_match(school_things, "^not.*[[:alpha:]]")

```


4. Practicing conditions by finding the expressions that start with "n" or "p". 
```{r}
str_match(school_things, "[n|p].*")
```

5. Grouping with conditions! Take the words that have two repeated vowels.

```{r}
str_match(school_things,"\\w*(a|e|i|o|u){2}.")
```



---

# Try it yourself!

As we said before, regular expressions are conventions to find and manipulate patterns in text data. Using the “str_match” function from the stringR package we can try out these conventions! 

*Note*: other stringR functions can be used with regular expressions, but we will not explain them because it is not the subject of our workshop. We invite you to go to the “String Manipulation with R” presentation to learn more about this package!

Before starting, some **reminders**:

1.  Practice makes perfect! You might not get the expression you want at first, but with patience and a little work you'll eventually get where you want. 

2. Read RegExps out loud to understand them! It is easier to remember them if you understand what they mean. For example, "slash d plus"(`"\\d+`) doesn't make sense until you remember that it means "give me one or more digits together"). (We thank Colin Quirk for this advice. You can see his tutorial, [_here_](https://www.youtube.com/watch?v=uJKpz9T7mAg). 

3. Remember that RegExp find patterns and then stop searching. Keep your repetition operators close and the brackets even closer!

*The story:* 
Imagine you decided to digitize one of your old notebooks. The following vectors were found on the first three pages.  

```{r}
shopping_list <-  c("tomato", "tangerine", "apple", "banana", "papaya", "tamarind")
contact_list <- c("lisa 444-4444-5555", "anna linkedin", "kevin’s number 344456632", "juan ?", "mary: linkedin") 
To_do_list <-c("shopping and banking", "laundry", "university stuff", "salsa and samba")

```

We will use these vectors to practice the key regular expressions and their combinations. But first, let's group them together to be able to remember them with more ease:   

1. Anchors and wild cards
  - `^` = Beginning of a string
  - `$` = End of a String
  - `.` = wild card
2. Repetitions
  - `+` = 1 or more
  - `*` = 0 or more
3. Letters that have meanings
  - `\d` = digits
  - `\w` = alphanumeric characters
  - `\s` = spaces
4. Brackets 
  - `[ ]` = Matching characters or brackets
  - `( )` = Grouping
  - `{ }` = Repeated 
5. Conditionals
  -  `|` = Either, or


**Exercise 1: Basic Combinations of**

Use first the `shopping_list` to practice anchors and wild cards. Remember to "read" the RegExp out loud! 

```{r}
shopping_list %>% str_match(".")

shopping_list %>% str_match("^t")
```

**Exercise 2:  **


---


# Where should you go to learn more

There are a bunch of cheatsheets out there. We recommend the one you can find on the [Stringr website](https://stringr.tidyverse.org/). If you want to practice using regular expressions, you can also visit [this website](https://regexone.com/).
 

---

# Sources {-}

1. [_R for Data Science_](https://r4ds.had.co.nz/strings.html)
2. [_Automated Data Collection with R. A Practical Guide to Web Scraping and Text Mining_](https://onlinelibrary.wiley.com/doi/book/)
3. [_Interactive Tutorial on Regular Expressions_](https://regexone.com/)
<<<<<<< HEAD
4. [_R Tutorial | Regular Expressions in R](https://www.youtube.com/watch?v=uJKpz9T7mAg)
5. [_Cheatsheet Stringr](https://stringr.tidyverse.org/)
=======
4. [_R Tutorial | Regular Expressions in R_](https://www.youtube.com/watch?v=uJKpz9T7mAg)
>>>>>>> 448646950747dd0d45b3b46ec305ba1983177c3c


